---
title: "Final Project"
author: "Linh Tran"
date: "5/7/2021"
output: 
   html_document:
     toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)

library(caret)
library(glmnet)
library(mlbench)
library(e1071)
library(kernlab)
library(pROC) #generate ROC curve and calculate AUC 
library(ROCR)
library(vip) #variable importance plot: global impact on different predictor
library(pdp)
library(randomForest)
library(ranger)
library(AppliedPredictiveModeling) # for visualization purpose
library(corrplot)
library(RColorBrewer)
library(RANN)
library(visdat)
library(mgcv)
library(DMwR2)
library(lime)
library(grid)
library(naniar)
library(knitr)
library(kableExtra)
library(rpart.plot)
library(rpart)
library(ROSE)
library(iml)
library(gridExtra)
library(cowplot)

knitr::opts_chunk$set(
  fig.width = 12,
  fig.height = 8,
  out.width = "90%"
)
```

# Import data

```{r data prep}
stroke_df = read.csv("./data/healthcare-dataset-stroke-data.csv")
# head(stroke_df)

```



# EDA

# EDA and Visualization

Distribution of stroke:

```{r}
dataplot10 = stroke_df %>% dplyr::count(stroke) 
dataplot1 = dataplot10 %>% mutate(ntotal=sum(dataplot10$n), perc= n/ntotal)
plot1= ggplot(dataplot1, aes(x="", y=perc*100, fill=as.factor(stroke), group=as.factor(stroke)))+theme_bw()+
  geom_bar(width = 1, stat = "identity") + theme_void() +
  labs(x=" ",y=" ", fill=" ") + 
  scale_fill_brewer(palette = "Dark2",labels = c("No stroke", "Stroke"))+
  geom_text( y=55, label="95.13 %", size=5)+geom_text(aes(label="4.87 %"),y=2.5, x=1.3, size=4)+
  coord_polar("y", start=0) + theme(legend.text=element_text(size=15))

plot1
```

We could see that only 4.87% of the 5110 individuals in the dataset suffered a stroke. \


```{r gender}
#genders

dataplot2=stroke_df %>% dplyr::count(stroke, gender) %>% spread(stroke, n)
names(dataplot2)=c("gender", "neg", "pos")

dataplot2 = dataplot2 %>% mutate(perc_gender=pos/(pos+neg))

plot2 = ggplot(dataplot2 %>% filter(gender!="Other"), aes(x=gender,
                        y=perc_gender*100, fill=as.factor(gender), 
                        group=as.factor(gender))) + theme_bw()+
  geom_bar(stat = "identity")+
  labs(title="Gender",x="",y="Probability of stroke (%)") + scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "none")+  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))

```

Smoking status:

```{r smoking status}

dataplot3_1=stroke_df %>% dplyr::count(stroke, smoking_status) %>% spread(stroke, n)
names(dataplot3_1)=c("smoking_status", "neg", "pos")

dataplot3_1 = dataplot3_1 %>% mutate(perc_smoke=pos/(pos+neg))

plot3 = ggplot(dataplot3_1, aes(x=smoking_status,
                        y=perc_smoke*100, fill=as.factor(smoking_status), 
                        group=as.factor(smoking_status))) + theme_bw()+
  geom_bar(stat = "identity")+
  labs(title="Smoking status",x=" ",y="Probability of stroke (%)") + scale_fill_brewer(palette = "Dark2") +
  scale_x_discrete(labels=c("formerly smoked" = "Formerly smoked", "never smoked" = "Never smoked", "smokes"="Smokes")) +
  theme(legend.position = "none")+  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))


```

People who identified as former smokers have the highest probability of having a stroke (~8%), followed by smokers and then people who never smoked.


```{r hypertension}
# hypertension

dataplot3_1a=stroke_df %>% dplyr::count(stroke, hypertension) %>% spread(stroke, n)
names(dataplot3_1a)=c("hypertension", "neg", "pos")

dataplot3_1a = dataplot3_1a %>% mutate(perc_hyp=pos/(pos+neg))

plot4 =ggplot(dataplot3_1a, aes(x=as.factor(hypertension) ,
                        y=perc_hyp*100, fill=as.factor(hypertension ), 
                        group=as.factor(hypertension ))) + theme_bw()+
  geom_bar(stat = "identity")+
  labs(title="Hypertension", x=" ",y="Probability of stroke (%)", fill=" ") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_discrete(breaks=c("0","1"), labels=c("0" = "No hypertension", "1" = "Hypertension")) + theme(legend.position = "none")+
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))



```


```{r heart disease}
#heart disease
dataplot3_1b=stroke_df %>% dplyr::count(stroke, heart_disease) %>% spread(stroke, n)
names(dataplot3_1b)=c("heart_disease", "neg", "pos")

dataplot3_1b = dataplot3_1b %>% mutate(perc_hd=pos/(pos+neg))

plot5 = ggplot(dataplot3_1b, aes(x=as.factor(heart_disease) ,
                         y=perc_hd*100, fill=as.factor(heart_disease ), 
                         group=as.factor(heart_disease ))) + theme_bw()+
  geom_bar(stat = "identity")+
  labs(title="Heart disease",x="", y="Probability of stroke (%)", fill="Heart disease") +
  scale_fill_brewer(palette = "Dark2") + 
  scale_x_discrete(breaks=c("0","1"), labels=c("0" = "No HD", "1" = "HD")) + theme(legend.position = "none")+
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))

```


```{r marital status}
#ever_married
dataplot3_1c=stroke_df %>% dplyr::count(stroke, ever_married) %>% spread(stroke, n)
names(dataplot3_1c)=c("ever_married", "neg", "pos")

dataplot3_1c = dataplot3_1c %>% mutate(perc_em=pos/(pos+neg))

plot6 = ggplot(dataplot3_1c, aes(x=ever_married ,
                         y=perc_em*100, fill=as.factor(ever_married ), 
                         group=as.factor(ever_married ))) + theme_bw()+
  geom_bar(stat = "identity")+
  labs(title="Ever married",x="", y="Probability of stroke (%)", fill=" ") +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "none")+ 
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))
```


```{r worktype}
# work type

dataplot3_1d= stroke_df %>% dplyr::count(stroke, work_type) %>% spread(stroke, n)
names(dataplot3_1d)=c("work_type", "neg", "pos")

dataplot3_1d = dataplot3_1d %>% mutate(perc_wt=pos/(pos+neg))

plot7=ggplot(dataplot3_1d %>% filter(work_type!="Never_worked"), aes(x=work_type ,
                         y=perc_wt*100, fill=as.factor(work_type ), 
                         group=as.factor(work_type ))) + theme_bw()+
  geom_bar(stat = "identity")+
  labs(title="Work type",x=" ",y="Probability of stroke (%)", fill=" ") +
  scale_fill_brewer(palette = "Dark2") + 
  scale_x_discrete(labels=c("children" = "Children", "Govt_job" = "Gov. Job")) +
  theme(legend.position = "none")+
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))
```


```{r}
#residence type

dataplot3_1e=stroke_df %>% dplyr::count(stroke, Residence_type) %>% spread(stroke, n)
names(dataplot3_1e)=c("Residence_type", "neg", "pos")

dataplot3_1e = dataplot3_1e %>% mutate(perc_rt=pos/(pos+neg))

plot8 = ggplot(dataplot3_1e, aes(x=Residence_type ,
                         y=perc_rt*100, fill=as.factor(Residence_type ), 
                         group=as.factor(Residence_type ))) + theme_bw()+
  geom_bar(stat = "identity")+
  labs(title="Residence type",x=" ", y="Probability of stroke (%)", fill=" ") +
  scale_fill_brewer(palette = "Dark2") +
  theme(legend.position = "none")+
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))
```

Categorical variables:

```{r figures of distribution of categorical outcomes}
#figures
allplotslist_1 <- align_plots(plot2, plot3, plot4, plot5, plot6, plot7, plot8, align = "hv")


grid_1=grid.arrange(allplotslist_1[[1]],allplotslist_1[[2]],
                  allplotslist_1[[3]],allplotslist_1[[4]],
                  allplotslist_1[[5]], allplotslist_1[[6]],nrow = 3)
```


Continuous variable:

```{r continuous variable}
#age
plot9 = 
  stroke_df %>% 
  ggplot() + 
  geom_density(aes(x=age  , group=as.factor(stroke),fill=as.factor(stroke)),
               size=1,alpha=0.5, adjust=2)  + 
  theme_bw()+
  ylab("Density")+ labs(fill=' ',x="Age") +   
  scale_fill_brewer(palette = "Dark2",labels = c("No stroke", "Stroke"))+
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))



# bmi
plot10 = 
  stroke_df %>% 
  ggplot() + 
  geom_density(aes(x=bmi, group=as.factor(stroke),fill=as.factor(stroke)),
               size=1,alpha=0.5, adjust=2)  + 
  theme_bw()+
  ylab("Density")+ labs(fill=' ',x="BMI") +   
  scale_fill_brewer(palette = "Dark2",labels = c("No stroke", "Stroke"))+
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))



# avg_glucose_level
plot11 = 
  stroke_df %>% 
  ggplot() + 
  geom_density(aes(x=avg_glucose_level  , group=as.factor(stroke),fill=as.factor(stroke)),
               size=1,alpha=0.5, adjust=2)  + 
  theme_bw()+
  ylab("Density")+ labs(fill=' ',x="Avg. glucose level") +   
  scale_fill_brewer(palette = "Dark2",labels = c("No stroke", "Stroke"))+
  theme(text = element_text(size=13.07,colour="black"))+
  theme(axis.text.x = element_text(colour="black",size=13.07))+
  theme(axis.text.y = element_text(colour="black",size=13.07))

#combine plots
plot9 + plot10 + plot11

```


**Comment**: From these plots we can see that:

Formerly smokers are more prone to suffer a stroke than smokers. It is odd, but I suspect that people, in general, do not quit smoking until they really need to quit. Therefore, maybe the formerly smokers are people who quit smoking because some health conditions which may increase their risk of suffering a stroke (it's just a brainstorm).

Self-employed are under higher risk of suffering a stroke than private and governament jobs.

Urban residents, males and people with hypertension or heart diasease are prone to suffer a stroke. In addition, people who have been married are also more likely to suffer a stroke than the single people.


## Change categorical variables to binary for model training

```{r}

stroke_df$stroke = as.factor(stroke_df$stroke)
stroke_df$gender = factor(stroke_df$gender) %>% as.numeric()
stroke_df$ever_married = factor(stroke_df$ever_married) %>% as.numeric()
stroke_df$work_type = factor(stroke_df$work_type) %>% as.numeric()
stroke_df$Residence_type = factor(stroke_df$Residence_type) %>% as.numeric()
stroke_df$smoking_status = factor(stroke_df$smoking_status) %>% as.numeric()
stroke_df$heart_disease = factor(stroke_df$heart_disease) %>% as.numeric()
stroke_df$hypertension = as.numeric(factor(stroke_df$hypertension))
stroke_df$work_type = as.factor(stroke_df$work_type) %>% as.numeric()
stroke_df$bmi = as.numeric(stroke_df$bmi)

stroke_df = stroke_df[, -1] %>% 
    mutate(stroke = recode(stroke, 
                           `0` = "No", 
                           `1` = "Yes"), 
           stroke = factor(stroke)) %>% 
    filter(gender < 3) 


summary(stroke_df)
vis_miss(stroke_df)

head(stroke_df)
```



## Partition the dataset

```{r}
set.seed(123)
trRow = createDataPartition(y = stroke_df$stroke, p = 0.7, list = F)
train.data = stroke_df[trRow, ]
test.data = stroke_df[-trRow, ] 

```

## Imputation with `preProcess() `

```{r}
knnImp = preProcess(train.data, method = "knnImpute", k = 3)
train.data = predict(knnImp, train.data)
vis_miss(train.data)
test.data = predict(knnImp,test.data)
vis_miss(test.data)
test.data$stroke
```


# Models





# Compare models