---
title: "Report draft"
author: "Linh Tran"
date: "3/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
set.seed(2021)
# load libraries
library(tidyverse) # metapackage of all tidyverse packages
library(naniar) # handling missing data
library(skimr) # quick overview over the dataset
library(caret) # ML toolkit
library(MLmetrics) # F1 Score
library(imbalance) # algorithms to deal with imbalanced datasets
library(gridExtra) # display plots in grids
library(patchwork) # arrange plots side by side
library(glmnet)
library(mlbench)  
library(pROC)   
library(pdp) 
library(vip) 
library(AppliedPredictiveModeling)
library(e1071)
library(MASS) 
library(klaR)
library(lares)
library(randomForest)

set.seed(88)

# custom plot size function
fig <- function(width, heigth){
     options(repr.plot.width = width, repr.plot.height = heigth)
}

## ggplot custom theme
theme_bigfont <- theme(plot.title = element_text(size=22),
                       axis.text.x= element_text(size=15),
                       axis.text.y= element_text(size=15), 
                       axis.title=element_text(size=18),
                       legend.text = element_text(size = 14))

# read data into R
stroke_data <- read_csv("data/healthcare-dataset-stroke-data.csv")

```



```{r}
# check the first few rows
head(stroke_data)

# summary of the data
summary(stroke_data)
```

```{r}
# how many "N/A" values are in my dataset per column?
miss_scan_count(data = stroke_data, search = list("N/A", "Unknown"))
```

```{r}
fig(15, 8)

stroke_data %>%
group_by(smoking_status) %>%
summarise(count = length(smoking_status)) %>%
mutate(smoking_status = factor(smoking_status)) %>%
ggplot(aes(x = fct_reorder(smoking_status, count), y = count, fill = factor(ifelse(smoking_status=="Unknown","Unknown","Known")))) +
geom_col() +
geom_text(aes(label = count, x = smoking_status, y = count), size = 6, hjust = 1.5) +
coord_flip() +
scale_fill_manual(values = c("Unknown" = "red", "Known" = "darkgrey")) +
labs(x = "smoking status") +
theme(legend.position = "none") +
theme_bigfont
```


```{r}
stroke_data_clean <- replace_with_na(data = stroke_data, replace = list(bmi = c("N/A"), smoking_status = c("Unknown"))) %>%
    # change bmi to numeric 
    mutate(bmi = as.numeric(bmi))

# check
summary(stroke_data_clean)
unique(stroke_data_clean$smoking_status)
```


```{r}
fig(15, 8)

# visualize the missing values
vis_miss(stroke_data_clean, cluster = TRUE) +
theme_bigfont
```


```{r}
fig(20, 30)

# create vector of column names with
cols <- stroke_data_clean %>% 
  dplyr::select(-id, -smoking_status) %>% 
  names()

vis_plots_list <- list()

for (i in 1:length(cols)) {
    vis_plots_list[[i]] <- stroke_data_clean %>% arrange_at(cols[i]) %>% vis_miss() + labs(title = paste0("Ordered by ", cols[i]))
}

n <- length(vis_plots_list)
nCol <- floor(sqrt(n))
do.call("grid.arrange", c(vis_plots_list, ncol = nCol))

```



```{r}
fig(10, 8)

# check distribution of bmi
ggplot(stroke_data_clean, aes(x = bmi)) +
geom_histogram() +
labs(title = "Distribution of BMI") +
theme_bigfont
```



```{r}
fig(10,8)

# impute median and bind shadow to evaluate imputation
stroke_data_imp <- bind_shadow(stroke_data_clean) %>% 
impute_median_at(.vars = c("bmi")) %>%
add_label_shadow()

# Explore the median values in bmi in the imputed dataset
ggplot(stroke_data_imp, 
       aes(x = bmi_NA, y = bmi)) + 
geom_boxplot() +
labs(title = "Comparison, no-missing vs. imputed values for BMI") +
theme_bigfont
```


```{r}
stroke_data_imp <- impute_median_at(stroke_data_clean, .vars = c("bmi"))
```


```{r}
fig(16,8)

p1 <- ggplot(stroke_data_imp, 
       aes(x = smoking_status, fill = smoking_status)) + 
geom_bar() +
labs(title = "Before filling in NA values in smoking_status") +
theme(legend.position = "none") +
theme_bigfont

# fill imputation based on previous unique value in "smoking_status" column
after <- stroke_data_imp %>% 
fill(smoking_status)
# mode imputation which leads to worse performance of models:
#mutate(across(c(smoking_status)), replace(., is.na(.), "never smoked"))

# Explore the median values in bmi in the imputed dataset
p2 <- ggplot(after, 
       aes(x = smoking_status, fill = smoking_status)) + 
geom_bar() +
labs(title = "After filling in NA values in smoking_status") +
theme(legend.position = "none") +
theme_bigfont

p1 + p2
```


```{r}
stroke_data_imp2 <- stroke_data_imp %>%
  fill(smoking_status) %>%
  mutate(across(c(hypertension, heart_disease), factor),
      across(where(is.character), as.factor),
      across(where(is.factor), as.numeric),
      stroke = factor(ifelse(stroke == 0, "no", "yes")))
```



```{r}
stroke_data_imp2 <- stroke_data_imp2 %>%
mutate(bmi = case_when(bmi < 18.5 ~ "underweight",
                      bmi >= 18.5 & bmi < 25 ~ "normal weight",
                      bmi >= 25 & bmi < 30 ~ "overweight",
                      bmi >= 30 ~ "obese"),
      bmi = factor(bmi, levels = c("underweight", "normal weight", "overweight", "obese"), order = TRUE))
```


```{r}
fig(10, 8)

# plot prop of people who had a stroke
stroke_data_imp2 %>%
dplyr::select(stroke) %>%
ggplot(aes(x = stroke)) +
geom_bar() +
theme_bigfont

# count how many people had a stroke and the prop
stroke_data_imp2 %>%
group_by(stroke) %>%
summarize(n = n()) %>%
mutate(prop = round(n / sum(n), 2))
```


Because we have imbalanced data, I decided to use oversampling method to increase the number of 'stroke' instances to balance the distribution of classes. 

```{r}

imbalanceRatio(as.data.frame(stroke_data_imp2), classAttr = "stroke")
```



```{r}
stroke_test <- stroke_data_imp2 %>%
mutate(
    stroke = as.character(stroke),
    across(where(is.factor), as.numeric),
    stroke = factor(stroke)
)

stroke_oversampled <- oversample(as.data.frame(stroke_test), classAttr = "stroke", ratio = 1, method = "MWMOTE")

head(stroke_oversampled)

stroke_oversampled %>%
  group_by(stroke) %>%
  summarize(n = n()) %>%
  mutate(prop = round(n / sum(n), 2))
```



```{r}
stroke_oversampled <- 
  stroke_oversampled %>% 
  dplyr::select(-id)


```


Train/Test split for oversampled data


```{r}
set.seed(2021)

trRow = createDataPartition(y = stroke_oversampled$stroke, p = 0.7, list = F)
train_oversamp = stroke_oversampled[trRow, ]
test_oversamp = stroke_oversampled[-trRow, ] 
```

Train/Test split for original data

```{r}
set.seed(2021)

trRow = createDataPartition(y = stroke_test$stroke, p = 0.7, list = F)
train_original = stroke_test[trRow, ]
test_original = stroke_test[-trRow, ] 

test_original_no_stroke = test_original %>% 
  dplyr::select(-id, - stroke)

```


# Model building

## GLMnet

```{r}
# custom train control
myControl <- trainControl(
  method = "cv", 
  number = 10,
  summaryFunction = twoClassSummary,
  classProbs = TRUE,
  verboseIter = TRUE
)

myGrid <- expand.grid(
    alpha = c(0,1),
    lambda = seq(0.00001, 1, length = 20)
)

set.seed(42)
glmnet_model <- train(
    stroke ~ .,
    train_oversamp,
    method = "glmnet",
    tuneGrid = myGrid,
    trControl = myControl
)

plot(glmnet_model)

max(glmnet_model[["results"]]$ROC)

mm_test <- test_oversamp %>% 
  dplyr::select(-stroke)

glmnet_pred <- predict(glmnet_model, newdata = mm_test) 

confusionMatrix(glmnet_pred, factor(test_oversamp[["stroke"]]), positive = "yes")

```



The best model using GLMnet is a ridge regression classifier. However, it has lower accuracy than the baseline model (0.78 vs 0.95), however it's recall (sensitivity) is of course larger than baseline (0.74) meaning it has at least some predictive power to classify true positives correctly.

Now redo for original data (before oversampling)
```{r}
glmnet_pred_original <- predict(glmnet_model, newdata = test_original_no_stroke) 

confusionMatrix(glmnet_pred_original, factor(test_original[["stroke"]]), positive = "yes")
```


Using original data (no oversampling)

```{r}
# custom train control
myControl <- trainControl(
  method = "cv", 
  number = 10,
  summaryFunction = twoClassSummary,
  classProbs = TRUE,
  verboseIter = TRUE
)

myGrid <- expand.grid(
    alpha = c(0,1),
    lambda = seq(0.00001, 1, length = 20)
)

set.seed(2021)
glmnet_model <- train(
    stroke ~ .,
    train_original,
    method = "glmnet",
    tuneGrid = myGrid,
    trControl = myControl
)

plot(glmnet_model)

max(glmnet_model[["results"]]$ROC)

mm_test <- test_original %>% 
  dplyr::select(-stroke)

glmnet_pred <- predict(glmnet_model, newdata = mm_test) 

confusionMatrix(glmnet_pred, factor(test_original[["stroke"]]), positive = "yes")
```


The values for accuracy improve a little for the oversampled data (0.7898 vs 0.7448). The sensitivity also improved for the oversampled data (0.8320 vs 0.68919).



## Random Forest

Original data

```{r}
rfGrid <- data.frame(
  .mtry = c(2,3,5,6),
  .splitrule = "gini",
  .min.node.size = 5
)

rfControl <- trainControl(
    method = "oob",
    number = 5,
    verboseIter = TRUE
)

rf_model_original <- train(
    stroke ~ .,
    train_original,
    method = "ranger",
    tuneLength = 3,
    tuneGrid = rfGrid,
    trControl = rfControl
)

rf_model_original

mm_test_original = test_original %>% 
  dplyr::select(-stroke)

rf_pred_original <- predict(rf_model_original, newdata = mm_test_original) 

confusionMatrix(rf_pred_original, factor(test_original[["stroke"]]), positive = "yes")


```


Oversampled data

```{r}
rfGrid <- data.frame(
  .mtry = c(2,3,5,6),
  .splitrule = "gini",
  .min.node.size = 5
)

rfControl <- trainControl(
    method = "oob",
    number = 5,
    verboseIter = TRUE
)

rf_model_oversamp <- train(
    stroke ~ .,
    train_oversamp,
    method = "ranger",
    tuneLength = 3,
    tuneGrid = rfGrid,
    trControl = rfControl
)

rf_model_oversamp

mm_test_oversamp = test_oversamp %>% 
  dplyr::select(-stroke)

rf_pred_oversamp <- predict(rf_model_oversamp, newdata = mm_test_oversamp) 

confusionMatrix(rf_pred_oversamp, factor(test_oversamp[["stroke"]]), positive = "yes")
```


Looks like the random forest model is great at classifying true negative cases, but performs poorly on classifying true positive cases which is what we are interested in (we want to detect people with stroke, so we can be confident in telling a patient they are at risk of stroke when we supply his/her information to the model).



