---
title: "P8106 Midterm"
author: "Linh Tran"
date: "3/27/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(caret)
library(glmnet)
library(mlbench)  
library(pROC)   
library(pdp) 
library(vip) 
library(AppliedPredictiveModeling)
library(e1071)
library(MASS) 
library(klaR)
library(lares)
library(naniar) #handling missing data
library(imbalance) #dealing with imbalanced datasets
library(gridExtra) #display plots in grids
library(patchwork)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Load the data

```{r}
stroke_df = read.csv("./data/healthcare-dataset-stroke-data.csv")

stroke_df$stroke = as.factor(stroke_df$stroke)
stroke_df$gender = as.factor(stroke_df$gender)
stroke_df$ever_married = as.factor(stroke_df$ever_married)
stroke_df$work_type = as.factor(stroke_df$work_type)
stroke_df$Residence_type = as.factor(stroke_df$Residence_type)
stroke_df$smoking_status = as.factor(stroke_df$smoking_status)
stroke_df$heart_disease = as.factor(stroke_df$heart_disease)
stroke_df$hypertension = as.factor(stroke_df$hypertension)
stroke_df$work_type = as.factor(stroke_df$work_type)
stroke_df$bmi = as.numeric(stroke_df$bmi)
stroke_df = stroke_df %>% 
    mutate(stroke = recode(stroke, 
                           `0` = "no stroke", 
                           `1` = "stroke")) %>% 
    dplyr::select(-id) %>% 
    filter(gender != "Other")
summary(stroke_df)
```

Missing values

```{r}
stroke_data = read.csv("./data/healthcare-dataset-stroke-data.csv")
miss_scan_count(data = stroke_data, search = list("N/A", "Unknown"))

```

There are 201 missing values in BMI and 1544 "Unknown" values in smoking_status. We can put this into perspective by using plot:

```{r}
stroke_df %>% 
  group_by(smoking_status) %>% 
  summarise(count = length(smoking_status)) %>% 
  ggplot(aes(x = fct_reorder(smoking_status, count), y = count)) + geom_col() +
  geom_text(aes(label = count, x = smoking_status, y = count))
```



```{r}
ggplot(stroke_df, aes(x = bmi)) +
geom_histogram() +
labs(title = "Distribution of BMI") 
```

The distribution is right-skewed. Because this is the only variable with missing data (at least the numerical variables) we can impute the median on the missing data.

```{r}

stroke_data_clean <- replace_with_na(data = stroke_data, replace = list(bmi = c("N/A"), 
                                                                        smoking_status = c("Unknown"))) %>%
    mutate(bmi = as.numeric(bmi))

# impute median and bind shadow to evaluate imputation
stroke_data_imp <- bind_shadow(stroke_data_clean) %>% 
impute_median_at(.vars = c("bmi")) %>%
add_label_shadow()

# Explore the median values in bmi in the imputed dataset
ggplot(stroke_data_imp, 
       aes(x = bmi_NA, y = bmi)) + 
geom_boxplot() +
labs(title = "Comparison, no-missing vs. imputed values for BMI") 


```


Next we can look at the proportion of people who have a stroke

```{r}
stroke_df_imp %>%
  dplyr::select(stroke) %>%
  ggplot(aes(x = stroke)) +
  geom_bar() 


# Count how many people who have a stroke
stroke_df_imp %>% 
  group_by(stroke) %>% 
  summarize (n = n()) %>% 
  mutate(prop = round(n/sum(n), 2)) %>% 
  knitr::kable()
```

We see that only 5% of all the people in the dataset had a stroke at some point. This means that our baseline dummy model has an accuracy of 95%. 
